---
title: "RNA-Seq significance analysis"
subtitle: "220221_A01275_0052_AHT7L3DRXY"
author: "Arnaud Muller"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:   
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(fig.width=16, fig.height=10) 
#knitr::opts_chunk$set(dev='png',dev.args=list(type='cairo'))
```


```{r loadlib}
suppressPackageStartupMessages(library("tidyverse"))
suppressPackageStartupMessages(library("stringr"))
suppressPackageStartupMessages(library("DESeq2"))
suppressPackageStartupMessages(library("pheatmap"))
suppressPackageStartupMessages(library("RColorBrewer"))
suppressPackageStartupMessages(library(FactoMineR))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library("viridis"))
suppressPackageStartupMessages(library("ggrepel"))
suppressPackageStartupMessages(library(biomaRt))
suppressPackageStartupMessages(library("DT"))
suppressPackageStartupMessages(library(ggforce))
suppressPackageStartupMessages(library(cowplot))
suppressPackageStartupMessages(library(UpSetR))
suppressPackageStartupMessages(library("sknifedatar"))
suppressPackageStartupMessages(library("xaringanExtra"))
suppressPackageStartupMessages(library("ggplotify")) 
suppressPackageStartupMessages(library(DEGreport))
suppressPackageStartupMessages(library("ComplexHeatmap"))
```

# File import

Initial data import stands in the following files:  

- **all.tsv**: raw counts 
- **sampleannotation.tsv**: metadata describing each sample, as seen below.

```{r import}
prefix <- "K:/arnaud/Projects/J_Phan_LIH/"
#prefix <- "./"

SampleAnnotation <- read_tsv(paste0(prefix,"sampleannotation.tsv"))

rawcounts <-  read_tsv(paste0(prefix,"all.tsv"))

datatable(SampleAnnotation,
          caption="Sample annotation", 
          extensions = 'Buttons', 
          options = list(dom = 'Bfrtip', buttons = list(I('colvis'),'copy', list(extend = 'collection',buttons = c('excel','csv','pdf'),text = 'Download')), autoWidth = F),rownames = FALSE)
```

```{r mart}
ensembl = useMart(biomart="ensembl", dataset="hsapiens_gene_ensembl")

genes <- rawcounts %>% pull(gene)

GeneAnnotation<- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol","gene_biotype"), filters = "ensembl_gene_id", values = genes, mart= ensembl)

```

# Distributions


```{r sanitary}
ddsrawcounts <- rawcounts %>% column_to_rownames("gene")
ddssampleannotation <- SampleAnnotation %>% column_to_rownames("sample")
ddsrawcounts <- ddsrawcounts[ , rownames(ddssampleannotation) ]
dds <- DESeqDataSetFromMatrix(countData=ddsrawcounts[ , rownames(ddssampleannotation) ], 
                              colData=ddssampleannotation, 
                              design=~condition)

dds <- dds[ rowSums(counts(dds)) > 10, ]

# Bar plot of lib size
colSums(assay(dds)) %>% as_tibble(rownames="sample") %>% 
  left_join(SampleAnnotation) %>% 
  ggplot()+
  geom_col(aes(x=sample, y=value, fill=condition), show.legend = F)+ 
  coord_flip()+ 
  facet_wrap(~condition,scales = "free_y") +
  theme_bw() +
  labs(title = "Barplot of library sizes", y = "Total read counts")+
  scale_y_continuous(expand = c(0, 0), limits = c(0, NA))+
  theme(text = element_text(size = 16, face = "bold"))

# Gene densities
ggplot(data=log2(counts(dds) +1) %>% as_tibble(rownames= "gene") %>% pivot_longer(-gene, names_to = "sample", values_to = "count") %>% left_join(SampleAnnotation))+
    geom_density(aes(count,color=sample, linetype=as.factor(time)),size=1.2)+
    labs(title = "Densities of raw log2 counts", x = "log2 counts + 1") +
    theme_bw() +
    theme(legend.position="bottom", legend.box = "vertical")+guides(color="none")+
    facet_wrap(~condition  ) +
    theme(text = element_text(size = 16, face = "bold"))

# Biotype plot
counts(dds)[rowSums(counts(dds)) >=  1000,] %>% 
    as_tibble(rownames= "gene") %>% 
    pivot_longer(-gene, names_to = "sample", values_to = "count") %>% 
    left_join(SampleAnnotation) %>% 
    left_join(GeneAnnotation, by=c("gene"="ensembl_gene_id")) %>% mutate(gene_biotype=as_factor(gene_biotype)) %>% 
    group_by(gene_biotype ) %>%
    ggplot()+
    geom_col(aes(x=sample, y=log2(count +1), fill=gene_biotype), position="fill")+
    scale_fill_manual(values=c(RColorBrewer::brewer.pal(5,'Paired'), RColorBrewer::brewer.pal(8,'Dark2'), RColorBrewer::brewer.pal(8,'Accent')))+    
    labs(x="Samples", y="Proportion of Biotype", fill="Biotype")+facet_wrap(~condition,scales = "free_y")+
    theme_bw() +
    coord_flip()+
    theme(text = element_text(size = 16, face = "bold")) +
    guides(fill=guide_legend(ncol=1))+
    scale_y_continuous(expand = c(0, 0), limits = c(0, NA))

```

# Clustering

```{r heatmap}
dds <- DESeq(dds)

vsd <- vst(dds)

sampleDistMatrix <- as.matrix(dist(t(assay(vsd))))

annot_colors <- list(time=c("30"="#DEF5E5FF","60"="#357BA2FF","90"="#40498EFF"))

pheatmap(sampleDistMatrix, 
         annotation_col=SampleAnnotation %>% dplyr::select(sample, time, condition) %>% mutate(time=as_factor(time), condition=as_factor(condition)) %>% column_to_rownames("sample"), 
         col=inferno(10), 
         annotation_colors = annot_colors)
```

```{r pca}
# calculate the variance for each gene
rv <- rowVars(assay(vsd))
# select the ntop genes by variance
select <- order(rv, decreasing=TRUE)[seq_len(min(500, length(rv)))]
pca <- prcomp(t(assay(vsd)[select,]))
# the contribution to the total variance for each component
percentVar <- pca$sdev^2 / sum( pca$sdev^2 )

pca$x %>% as_tibble(rownames = "sample") %>% 
  left_join(SampleAnnotation) %>% 
  ggplot(aes(PC1, PC2, color=condition, shape=as.factor(time))) +geom_point(size = 4) +
    labs(title = "PCA plot") + xlab("PC1 (53.1%)") + ylab("PC2 (23%)")+
    theme_bw() +
  theme(text = element_text(size = 16, face = "bold")) +
    theme(legend.position="bottom", legend.box = "vertical")
percentVar %>% as_tibble(rownames="PC")%>%
    ggplot()+
    geom_col(aes(x=fct_reorder(PC, -value), y=100 * value))+
    theme_bw() +
    theme(text = element_text(size = 16, face = "bold")) +
    scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +labs(title="variance explained by each PC",x="Principal Component", y="Percentage")

# Plot all first 5 PCs
ggplot(pca$x %>% as_tibble(rownames = "sample") %>%  left_join(SampleAnnotation) ) + 
    geom_point(alpha = 0.9,  size = 3, aes(x = .panel_x, y = .panel_y, color=condition, shape=as.factor(time))) +  geom_autodensity(aes(x = .panel_x, y = .panel_y), fill="grey60", alpha=.8) +
    facet_matrix(vars(PC1:PC5), layer.diag = 2, 
                 grid.y.diag = FALSE)+labs(title = "PCA plot") +
    theme_bw() +
    theme(text = element_text(size = 16, face = "bold")) +
    theme(legend.position="bottom", legend.box = "vertical")
```

# Results

For each comparison, you can find the box plot of top20 significant genes and a volcano plot.  
A table with *fold-change* and *adjusted pValue* is as well available for each comparison. 
The comparisons of interest are the folowing:  


```{r uberbasicmodel, cache=TRUE, cache.lazy = FALSE}
SampleAnnotation <- SampleAnnotation %>% unite("condition_time", condition, time, remove=F)
ddssampleannotation <- SampleAnnotation %>% column_to_rownames("sample")

# coldata <- ddssampleannotation %>% filter(condition_time %in% c("PA_30", "PB_30"))%>% dplyr::select(condition_time)
# countdata <- ddsrawcounts[,rownames(coldata)]
# design <- model.matrix(~  condition_time, coldata )
# dds <- DESeqDataSetFromMatrix(countData=countdata, 
#                               colData=coldata, 
#                               design= design)
# res <- results(dds)

fu_DESeq = function(key0,key1){
  coldata <- ddssampleannotation %>% filter(condition_time %in% c(key0,key1))%>% dplyr::select(condition_time)
  countdata <- ddsrawcounts[,rownames(coldata)]
  design <- model.matrix(~  condition_time, coldata )
  dds <- DESeqDataSetFromMatrix(countData=countdata, 
                              colData=coldata, 
                              design= design)
  dds <- estimateSizeFactors(dds)

  nc <- counts(dds, normalized=TRUE)
  filter <- rowSums(nc >= 1) >= 5 # this is to filter at least 1 (normalized) counts in 5 samples
  dds <- dds[filter,]

  dds <- DESeq(dds)

  #res <- results(dds)

  return(dds)
}


comparisons <- tibble::tribble(
                               ~group1, ~group2,
                               "PA_30", "PB_30",
                               "PA_60", "PB_60",
                               "PA_90", "PB_90",
                               "PA_30", "WTA_30",
                               "PA_60", "WTA_60",
                               "PA_90", "WTA_90",
                               "PB_30", "WTB_30",
                               "PB_60", "WTB_60",
                               "PB_90", "WTB_90",
                               "WTA_30", "WTB_30",
                               "WTA_60", "WTB_60",
                               "WTA_90", "WTB_90",
                 )
comparisons

AllResults <- comparisons %>% #filter(group1 == "WTA_30") %>%
  mutate(objDESeq=map2(group1, group2, ~fu_DESeq(.x, .y))) %>% 
  mutate(resDESeq=map(objDESeq, ~results(.x)%>% as_tibble(rownames = "gene"))) %>% 
  mutate(resDESeq=map(resDESeq, ~arrange(.x, pvalue))) %>% # sort by significance
  mutate(countsDESeq=map(objDESeq, ~counts(.x, normalized = T)%>% as_tibble(rownames = "gene"))) %>% 
  mutate(plotbox=pmap(list(resDESeq, countsDESeq,group1,group2), ~ggplot(data=..2 %>% filter(gene %in% pull(top_n(..1,-20, padj), gene ))  %>% pivot_longer(-gene, names_to = "sample", values_to = "count") %>% left_join(SampleAnnotation) %>% left_join(GeneAnnotation, by=c("gene"="ensembl_gene_id")), aes(x=condition_time, y=log2(count +1))) + 
                        geom_boxplot(outlier.shape = NA, notch=F) + 
                        geom_jitter(aes(color=as.factor(replicate)), width = 0.2, size=3.5)+
                        facet_wrap(~hgnc_symbol, scales="free_y") +
                        theme_bw() +
                        theme(text = element_text(size = 16, face = "bold"), axis.text.x=element_text(angle=45,hjust=1))+
                        theme(legend.position="bottom")+
                        labs(subtitle=paste0(..4," versus ",..3), title = "Top 20 significant genes")
  )) %>% 
  mutate(plotvolcano=pmap(list(resDESeq,group1,group2), ~ggplot(data=..1 %>% left_join(GeneAnnotation, by=c("gene"="ensembl_gene_id")), aes(x=log2FoldChange, y=-log10(padj), label=hgnc_symbol)) +
                            geom_point() + 
                            theme_minimal() +
                            theme(text = element_text(size = 16, face = "bold"))+
                            geom_text_repel(data=. %>% top_n(-20, padj)) +
                            #scale_color_manual(values=c("blue", "black", "red")) +
                            geom_vline(xintercept=c(-0.6, 0.6), col="red", linetype="dashed") +
                            geom_hline(yintercept=-log10(0.05), col="red")+
                            labs(subtitle=paste0(..3," versus ",..2), title = "Volcano plot")
  )) %>% 
  mutate(heatmap=pmap(list(resDESeq, objDESeq,group1,group2), ~pheatmap(..2 %>% vst() %>% assay() %>% as_tibble(rownames = "gene") %>% filter(gene %in% pull(top_n(..1,-20, padj), gene )) %>% column_to_rownames("gene"), show_rownames=FALSE, scale = "row", silent=T ))) %>% 
  mutate(plotall=pmap(list(plotbox, plotvolcano, heatmap), ~ggdraw() +
    draw_plot(..1, 0, 0, 0.5, 1) +
    draw_plot(..2, 0.5, 0.5, .5, .5) +
    draw_plot(as.ggplot(..3), 0.5, 0, 0.5, .5) +
    draw_plot_label(c("A", "B", "C"), c(0, 0.5, 0.5), c(1, 1, 0.5), size = 15)))
  #mutate(plotall=map2(plotbox, plotvolcano, ~plot_grid(.x,.y, ncol = 2)))


```

```{r plotresults}
#AllResults$plotall 
```

```{r panelset}
xaringanExtra::use_panelset()
AllResults <- AllResults %>% mutate(comparison=paste0(group1,"vs",group2))
```

`r library(sknifedatar); automagic_tabs(input_data = AllResults, panel_name = "comparison", .output = "plotall")`

The number of significant genes (adjPval < 0.01) per comparison is the following:  

```{r export}
# Number SDE genes for each group
AllResults %>% dplyr::select(group1, group2,resDESeq) %>% unite("group1_group2",group1,group2) %>% unnest(resDESeq) %>% group_by(group1_group2) %>% filter(padj<0.01) %>%  summarise(n = n())

AllResults%>% dplyr::select(group1, group2,resDESeq) %>% unite("group1_group2",group1,group2, sep="_versus_") %>% unnest(resDESeq) %>% filter(padj<0.01) %>% dplyr::select(group1_group2, padj, gene) %>% mutate(padj= case_when(!(is.na(padj)) ~ 1, TRUE ~ 0)) %>% pivot_wider(names_from = group1_group2, values_from = padj, values_fill = 0) %>% column_to_rownames("gene") %>% upset(., nsets = 12, number.angles = 30, point.size = 3.5, line.size = 2, mainbar.y.label = "Genre Intersections", sets.x.label = "SDE genes")

# Export All results files
#AllResults%>% #filter(group1 == "WTA_30") %>%
#  unite(comparison, c(group1, group2), remove=F, sep="_versus_") %>% 
#  mutate(walk2(resDESeq, comparison, ~write_tsv(.x, file=paste0(.y,".txt"))))
  
```
Here are the global comparisons (all time points merged together).  

```{r allTP}
# Patient A
coldata <- ddssampleannotation %>% filter(patient %in% c("A"))%>% dplyr::select(condition)
countdata <- ddsrawcounts[,rownames(coldata)]
design <- model.matrix(~  condition, coldata )
dds <- DESeqDataSetFromMatrix(countData=countdata, 
                              colData=coldata, 
                              design= design)
dds <- estimateSizeFactors(dds)

nc <- counts(dds, normalized=TRUE)
filter <- rowSums(nc >= 1) >= 5 # this is to filter at least 1 (normalized) counts in 5 samples
dds <- dds[filter,]

dds <- DESeq(dds)

resAllA <- results(dds)
resAllA <- resAllA%>% as_tibble(rownames = "gene") %>% arrange( pvalue)
write_tsv(resAllA, file="./PA_All_versus_WTA_All.txt")

# Patient B
coldata <- ddssampleannotation %>% filter(patient %in% c("B"))%>% dplyr::select(condition)
countdata <- ddsrawcounts[,rownames(coldata)]
design <- model.matrix(~  condition, coldata )
dds <- DESeqDataSetFromMatrix(countData=countdata, 
                              colData=coldata, 
                              design= design)
dds <- estimateSizeFactors(dds)

nc <- counts(dds, normalized=TRUE)
filter <- rowSums(nc >= 1) >= 5 # this is to filter at least 1 (normalized) counts in 5 samples
dds <- dds[filter,]

dds <- DESeq(dds)

resAllB <- results(dds)
resAllB <- resAllB%>% as_tibble(rownames = "gene") %>% arrange( pvalue)
write_tsv(resAllB, file="./PB_All_versus_WTB_All.txt")

```

# Advanced model

Here we propose to adjust the model in order to fit better with the experimental design:  
  
This is the design considered: `~ condition + time + condition:time, coldata`  
either for PA + WTA samples and for PB + WTB samples.

## PA and WTA samples

```{r advancedmodel, eval=TRUE}
coldata <-
  ddssampleannotation %>% filter(condition %in% c("PA", "WTA")) %>% dplyr::select(condition, time, replicate, condition_time) %>% mutate(
    condition = as_factor(condition),
    time = as_factor(time),
    replicate = as_factor(replicate)
  ) %>% unite(id, c("condition_time", "replicate"), remove = F) %>% dplyr::select(-condition_time)

countdata <- ddsrawcounts[, rownames(coldata)]
design <- model.matrix( ~ condition + time + condition:time, coldata)
ddsTCA <- DESeqDataSetFromMatrix(countData = countdata,
                                colData = coldata,
                                design = design)
ddsTCA <- estimateSizeFactors(ddsTCA)

nc <- counts(ddsTCA, normalized = TRUE)
filter <-
  rowSums(nc >= 1) >= 5 # this is to filter at least 1 (normalized) counts in 5 samples
ddsTCA <- ddsTCA[filter, ]

#ddsTCA <- DESeq(ddsTCA)
ddsTCA <- DESeqDataSet(ddsTCA, ~ condition + time + condition:time)
ddsTCA <- DESeq(ddsTCA, test = "LRT", reduced = ~ condition + time)
resTCA <- results(ddsTCA)

# top <- plotCounts(ddsTCA, which.min(resTCA$padj), intgroup = c("time","condition"), returnData = TRUE)
# top$time <- as.numeric(as.character(top$time))
# ggplot(top,
#   aes(x = time, y = count, color = condition, group = condition)) + 
#   geom_point() + stat_summary(fun.y=mean, geom="line") +
#   scale_y_log10()


res30 <- results(ddsTCA, name="conditionWTA.time60", test="Wald")
res30[which.min(resTCA$padj),]
betas <- coef(ddsTCA)
#colnames(betas)

topGenes <- head(order(resTCA$padj),20)
mat <- betas[topGenes, -c(1,2)]
thr <- 3 
mat[mat < -thr] <- -thr
mat[mat > thr] <- thr
#pheatmap(mat, breaks=seq(from=-thr, to=thr, length=101), cluster_col=FALSE)

top20 <- resTCA %>%as_tibble(rownames = "gene") %>% filter(padj<0.01) %>% top_n(-20, padj) %>% pull(gene)
counts(ddsTCA, normalized = T) %>% as_tibble(rownames = "gene")%>% filter(gene %in% top20)%>% pivot_longer(-gene, names_to = "sample", values_to = "count") %>% left_join(SampleAnnotation) %>% left_join(GeneAnnotation, by=c("gene"="ensembl_gene_id")) %>% 
  ggplot(aes(x=time, y=log2(count +1), color = condition, group = condition)) +geom_point() + stat_summary(fun.y=mean, geom="line", size=1.5) +
facet_wrap(~hgnc_symbol, scales="free_y") +
  theme_bw() +theme(text = element_text(size = 16, face = "bold")) +theme(legend.position="bottom")+labs(title = "Top 20 significant genes")

# phdata <- counts(ddsTC, normalized = T) %>% as_tibble(rownames = "gene")%>% filter(gene %in% top20)%>% pivot_longer(-gene, names_to = "sample", values_to = "count") %>% left_join(SampleAnnotation) %>% left_join(GeneAnnotation, by=c("gene"="ensembl_gene_id")) %>% dplyr::select(hgnc_symbol, sample, count) %>% pivot_wider(names_from = sample, values_from = count) %>% column_to_rownames("hgnc_symbol") %>% as.matrix()
# pheatmap(log2(phdata +1), 
#          annotation_col=SampleAnnotation %>% dplyr::select(sample, time, condition) %>% mutate(time=as_factor(time), condition=as_factor(condition)) %>% column_to_rownames("sample"), 
#          col=inferno(10), 
#          annotation_colors = annot_colors)

```

## PB and WTB samples

```{r advancedmodelB}
coldata <- ddssampleannotation %>% filter(condition %in% c("PB", "WTB")) %>% dplyr::select(condition, time, replicate, condition_time) %>% mutate(condition=as_factor(condition), time=as_factor(time), replicate=as_factor(replicate)) %>% unite(id, c("condition_time","replicate"), remove=F) %>% dplyr::select(-condition_time)

countdata <- ddsrawcounts[,rownames(coldata)]
design <- model.matrix(~ condition + time + condition:time, coldata )
ddsTCB <- DESeqDataSetFromMatrix(countData=countdata, 
                            colData=coldata, 
                            design= design)
ddsTCB <- estimateSizeFactors(ddsTCB)

nc <- counts(ddsTCB, normalized=TRUE)
filter <- rowSums(nc >= 1) >= 5 # this is to filter at least 1 (normalized) counts in 5 samples
ddsTCB <- ddsTCB[filter,]
  #ddsTC <- DESeq(ddsTC)
ddsTCB <- DESeqDataSet(ddsTCB, ~ condition + time + condition:time)
ddsTCB <- DESeq(ddsTCB, test="LRT", reduced = ~ condition + time)
resTCB <- results(ddsTCB)

# top <- plotCounts(ddsTC, which.min(resTC$padj), intgroup = c("time","condition"), returnData = TRUE)
# top$time <- as.numeric(as.character(top$time))
# ggplot(top,
#   aes(x = time, y = count, color = condition, group = condition)) + 
#   geom_point() + stat_summary(fun.y=mean, geom="line") +
#   scale_y_log10()


res30B <- results(ddsTCB, name="conditionWTB.time60", test="Wald")
res30B[which.min(resTCB$padj),]
betas <- coef(ddsTCB)
#colnames(betas)

topGenes <- head(order(resTCB$padj),20)
mat <- betas[topGenes, -c(1,2)]
thr <- 3 
mat[mat < -thr] <- -thr
mat[mat > thr] <- thr
#pheatmap(mat, breaks=seq(from=-thr, to=thr, length=101), cluster_col=FALSE)

top20 <- resTCB %>%as_tibble(rownames = "gene") %>% filter(padj<0.01) %>% top_n(-20, padj) %>% pull(gene)
counts(ddsTCB, normalized = T) %>% as_tibble(rownames = "gene")%>% filter(gene %in% top20)%>% pivot_longer(-gene, names_to = "sample", values_to = "count") %>% left_join(SampleAnnotation) %>% left_join(GeneAnnotation, by=c("gene"="ensembl_gene_id")) %>% 
  ggplot(aes(x=time, y=log2(count +1), color = condition, group = condition)) +geom_point() + stat_summary(fun.y=mean, geom="line", size=1.5) +
facet_wrap(~hgnc_symbol, scales="free_y") +
  theme_bw() +theme(text = element_text(size = 16, face = "bold")) +theme(legend.position="bottom")+labs(title = "Top 20 significant genes")
```

# Profile analysis

Among the results obtained using the advanced model, several genes share the same expression profile among time.  
These defferent generic profiles can be clustered together.

## PB and WTB samples

```{r profileA}
ma = assay(rlog(ddsTCA))[resTCA %>% as_tibble(rownames = "gene") %>% filter(padj<0.01) %>% pull(gene),]
design <- as.data.frame(colData(ddsTCA))
resPatternA <- degPatterns(ma, design, time = "time", col="condition")

ggplot(resPatternA[["normalized"]],
       aes(time, value, color = condition)) +
    geom_boxplot() +
    geom_point(position = position_jitterdodge(dodge.width = 0.9)) +
    # change the method to make it smoother
    geom_smooth(aes(group=condition), method = "loess")+facet_wrap(~cluster)

# phdata <- counts(ddsTC, normalized = T) %>% as_tibble(rownames = "gene")%>% filter(gene %in% top20)%>% pivot_longer(-gene, names_to = "sample", values_to = "count") %>% left_join(SampleAnnotation) %>% left_join(GeneAnnotation, by=c("gene"="ensembl_gene_id")) %>% dplyr::select(hgnc_symbol, sample, count) %>% pivot_wider(names_from = sample, values_from = count) %>% column_to_rownames("hgnc_symbol") %>% as.matrix()
# pheatmap(log2(phdata +1), 
#          annotation_col=SampleAnnotation %>% dplyr::select(sample, time, condition) %>% mutate(time=as_factor(time), condition=as_factor(condition)) %>% column_to_rownames("sample"), 
#          col=inferno(10), 
#          annotation_colors = annot_colors)

```

## PB and WTB samples

```{r profileB}
ma = assay(rlog(ddsTCB))[resTCB %>% as_tibble(rownames = "gene") %>% filter(padj<0.01) %>% pull(gene),]
design <- as.data.frame(colData(ddsTCB))
resPatternB <- degPatterns(ma, design, time = "time", col="condition")
```

```{r complexheatmap}
hmdataA <- t(scale(t(ma)))
 myCol <- colorRampPalette(c('dodgerblue', 'black', 'yellow'))(100)
  myBreaks <- seq(-3, 3, length.out = 100)
  
column_ha = HeatmapAnnotation(time= colnames(hmdataA) %>% as_tibble() %>% dplyr::select("sample" = value) %>% left_join(SampleAnnotation) %>% pull(time))

#Heatmap(hmdataA, name = "mat", top_annotation = column_ha)
order <- colnames(hmdataA) %>% as_tibble() %>% dplyr::select("sample" = value) %>% left_join(SampleAnnotation) %>% pull(sample)
#order <- rep(c(rep(c("30"), 5),c(rep(c("60"), 5) ),c(rep(c("90"), 5) )),2)

#Heatmap(hmdataA, name = "mat", top_annotation = column_ha, column_order = order, row_km = 6)
Heatmap(hmdataA, name = "mat", top_annotation = column_ha, column_order = order, row_km = 4, column_split =rep(c(rep(c("30"), 5),c(rep(c("60"), 5) ),c(rep(c("90"), 5) )),2), show_row_names=F)
```

# Session info

```{r sessioninfo, echo = TRUE,message=TRUE,warning=FALSE,cache=TRUE}
sessionInfo()
```
